---
title: "Exploratory Data Analysis of California's Housing Price: Insights and Trends"
author: "Nam Nguyen (USF)"
date: "October, 2024"
output: 
  html_document:
    theme: yeti
runtime: shiny
---

```{r setup, include = FALSE}
source("main.R")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## **Introduction**

This project aims to perform *Exploratory Data Analysis* and *Linear Regression* with the **California House Price** dataset.

Data source: [Link](https://www.kaggle.com/datasets/shibumohapatra/house-price)

Github file: [Link](https://github.com/nnamnguyenn/Exploratory-Data-Analysis-of-California-s-Housing-Price-Insights-and-Trends.git)

Library used: `readr`, `psych`, `corrplot`, `tidyverse`, `ggplot2`, `knitr`, `plotly`, `ggthemes`, `DT`, `DataExplorer`

## California House Price Dataset

```{r}
df |> head() |> kable() 
```

## **Summary Statistics Table**

### Overview

```{r,fig.show = "hold"}
df |> DataExplorer::introduce() |> knitr::kable()
```

### Continuous Data

```{r}
df |> psych::describe() |> round(2) |> kable()

```

### Categorical Data

```{r, , fig.width= 5, fig.height= 5, fig.align='center'}
df |> DataExplorer::plot_bar(theme_config = theme_bw()) 

```

## **Distribution**

```{r, echo = FALSE}
selectInput("display_value", label = "View: ",
            choices = setdiff(colnames(df), c("longitude", "latitude", "ocean_proximity", "total_bedrooms")), selected = 'median_house_value')
selectInput("break_down", label = "Distribution by Ocean Proximity: ",
            choices = c(TRUE, FALSE), selected = FALSE)

```

```{r, fig.show='hold', echo = FALSE}

renderPlotly(distribution_graph(input$display_value, input$break_down))
renderPlot(box_graph(input$display_value))

```

## **Correlation**

### Matrix

```{r, fig.show = 'hold'}
df |> select(!c(longitude, latitude, ocean_proximity, total_bedrooms)) |> cor() |> round(3) |> kable()
```

### Heatmap

```{r, fig.show = 'hold', fig.align='center'}
df |> select(!c(longitude, latitude, ocean_proximity, total_bedrooms)) |> cor() |> corrplot(method = 'color', order = 'AOE', type = "lower", tl.col = "black", tl.srt = 0.00001, tl.cex = 0.75)

```

## **Pair Plot**

```{r, fig.width= 10, fig.height= 10}
df |> select(!c(longitude, latitude, ocean_proximity, total_bedrooms)) |> slice_sample(prop = .05) |> pairs( panel = function(x, y) {
  points(x, y, col = "blue")  
  abline(lm(y ~ x), col = "red")})
```

## **Linear Regression (Median House Value \~ Median Income)**

```{r, eval=FALSE}
lm(median_house_value ~ median_income, data = df) |> summary() |> print()
```

```{r, echo = FALSE}
linearReg <- lm(median_house_value ~ median_income, data = df) 
linearReg |> summary() |> print()
```

### ANOVA test

```{r, eval = FALSE}
anova(linearReg)
```

```{r, echo = FALSE}
anova(linearReg)
```

### Model Evaluation Plot

```{r, echo = FALSE}
plot(linearReg)
```

## **Data Source**

```{r, echo = FALSE, fig.align='center'}
renderDT({
  datatable(df)
})
```
